Perturbing SU($n$)-symmetric interactions

### 2020-01-31 collective_sun/sun_perturbation.tex
add diagrammatic calculation of single- and two-body operator products start calculation of three two-body Ising-like operators,   which will be necessary to compute matrix elements of a perturbation in the asymmetric manifolds

% for drawing set (Venn) diagrams
\usepackage{tikz}
\tikzset{
  baseline = (current bounding box.center)
}
\usetikzlibrary{math}

@@

We wish to determine the effective dynamics induced on the
ground-state manifold $\M_0$ of $H_0$ by weak single- and two-body
perturbations of the form
\begin{align}
  \V_1 \equiv \sum_X \sum_p v_{Xp} X_p = \sum_X \v v_X\c\v X,
  &&
  \V_2 \equiv \sum_O \sum_{p<q} w_{Opq} O_{pq}
  = \sum_O \v{\m w}_O\c \v{\m O}
  \label{eq:perturbations}
\end{align}
where $p,q$ index individual spins; $v_{Xp}$ and $w_{Opq}$ are real
numbers; $X$ is a trace-zero single-body operator on $\H_n$; $O$ is a
trace-zero two-body operator on $\H_n\otimes\H_n$ that obeys pair-wise
permutational symmetry, i.e.~$O_{pq}=O_{qp}$; $\v v_X$ and $\v X$ are
$N$-component vectors defined by
\begin{align}
  \v v_X \equiv \sum_p v_{Xp} \ket{p},
  &&
  \v X \equiv \sum_p X_p \ket{p};
\end{align}
and finally $\v{\m w_O}$ and $\v{\m O}$ are ${N \choose 2}$-component
vectors defined by
\begin{align}
  \v{\m w}_O \equiv \sum_{p<q} w_{Opq} \ket{\set{p,q}},
  &&
  \v{\m O} \equiv \sum_{p<q} O_{pq} \ket{\set{p,q}},
\end{align}
where the basis vectors $\ket{\set{p,q}}$ are labeled by a choice of
two distinct spins $p,q$.

@@

If we {\it construct} a perturbation $\V_1^\Delta$ with coefficients
$\v v_X^\Delta$ that satisfy \eqref{eq:cond_1} with eigenvalue
$\Delta$, then
\begin{align}
  H_0 \V_1^\Delta \ket\psi = \p{E_0 + \Delta} \V_1^\Delta \ket\psi.
\end{align}
Interestingly, the energy of the state $\V_1^\Delta\ket\psi$ depends
only on the coefficients $\v v_X^\Delta$, and is entirely independent
of the state $\ket\psi$ or the choice of a trace-zero single-body
operators $X$ used to build $\V_1^\Delta$.  Finding operators
$\V_1^\Delta$ that generate eigenvectors of the interaction
Hamiltonian $H_0$ when they act on fully symmetric states $\ket\psi$
thus reduces to finding eigenvectors of the interaction matrix
$\m h-\diag\v h$.

@@

We now return to the task of computing the second-order effective
Hamiltonian $H_1^{(2)}$.  Any coefficient vector $\v v_X$ can be
expanded into its projections $\v v_X^\Delta$ onto the eigenspace of
vectors satisfying \eqref{eq:cond_1} with eigenvalue $\Delta$,
i.e.~$\v v_X = \sum_\Delta \v v_X^\Delta$, which also allows us to
expand
\begin{align}
  \V_1 = \sum_\Delta \V_1^\Delta,
  &&
  \V_1^\Delta \equiv \sum_X \v v_X^\Delta \c \v X,
\end{align}
where each operator $\V_1^\Delta$ generates a state with definite
interaction energy $\Delta$ above that of the fully symmetric manifold
$\M_0$.  We can therefore simplify
\begin{align}
  H_1^{(2)}
  = - \P_0 \V_1 \E \V_1 \P_0
  = - \sum_{\Delta\ne0} \f1{\Delta}
  \P_0 \V_1^\Delta \P_\Delta \V_1^\Delta \P_0
  = - \sum_{\Delta\ne0} \f1{\Delta} \P_0 \p{\V_1^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_1^\Delta}^2 \P_0$ is worked out in
Appendix \ref{sec:PXYP}.  In total, we find
\begin{align}
  H_1^{(2)}
  \EQFS \f1{N\p{N-1}} \sum_{X,Y} \sum_{\Delta\ne0}
  \f{\v v_X^\Delta\c\v v_Y^\Delta}{\Delta}
  \p{\col{X}\,\col{Y} - N \col{XY}},
  \label{eq:H_1_2}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$; $\v v_X^\Delta$ is the projection of
$\v v_X$ onto the $\Delta$-eigenspace of the matrix $\m h-\diag\v h$;
and for any single-body operator $Z$, we define the collective operator
\begin{align}
  \col{Z} \equiv \sum_p Z_p.
\end{align}

@@

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
\eqref{eq:perturbations} is simply the restriction of $\V_2$ onto
$\M_0$:
\begin{align}
  H_2^{(1)} = \sum_O \sum_{p<q} w_{Opq} \P_0 O_{pq} \P_0
  \EQFS \sum_O \sum_{p<q} w_{Opq} O_{1,2}
  \EQFS \sum_O \bar w_O \col{O},
\end{align}
where
\begin{align}
  \bar w_O \equiv {N \choose 2}^{-1} \sum_{p<q} w_{Opq},
  &&
  \col{O} \equiv \sum_{p<q} O_{pq},
\end{align}
are an average of $w_{Opq}$ and a collective version of the two-body
operator $O_{pq}$.  If the two-body operator $O=X\otimes Y$, then
\begin{align}
  H_2^{(1)} \EQFS \f12 \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
  \p{\col{X}\,\col{Y} - \col{XY}}.
\end{align}
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_VM_psi})
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \sum_O \sum_{p<q} \sp{\bar h_{pq} w_{Opq}
    + \sum_k \p{h_{kp} w_{Okq} w_{Okq} + h_{kq} w_{Opk}}}
  O_{pq} \ket\psi,
  \label{eq:H_V2_psi}
\end{align}
where to keep notation compact, we define
\begin{align}
  \bar h_{pq} \equiv 2 h_{pq} - h_p - h_q,
  &&
  h_p \equiv \sum_q h_{pq},
\end{align}
and we define $w_{Opq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$.  We thus find that the
vector $\V_2^\Delta\ket\psi$ (indexed by $\Delta$) is an eigenvector
of the interaction Hamiltonian $H_0$ if its coefficients
$\v{\m w}_O^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \check{\m h} \c \v{\m w} = \Delta \v{\m w},
  \label{eq:cond_2}
\end{align}
where
\begin{align}
  \check{\m h} \equiv \sum_{p<q} \bar h_{pq} \op{\set{p,q}}
  + \sum_{p<q} \ket{\set{p,q}} \sum_{k\notin\set{p,q}}
  \p{h_{kp} \bra{\set{k,q}} + h_{kq} \bra{\set{p,k}}}.
  \label{eq:h_super_mat}
\end{align}
Solving the general eigenvalue problem in \eqref{eq:cond_2} requires
diagonalizing the ${N \choose 2}\times{N \choose 2}\sim N^2\times N^2$
matrix $\check{\m h}$.  When both he interaction Hamiltonian $H_0$
translationally invariant, however, this eigenvalue problem can be
reduced to that of diagonalizing an $\sim N\times N$ matrix; we
discuss this reduction in Appendix \ref{sec:trans_inv}.

@@

In this section, we simplify the products of the form
$H_0\V_M\ket\psi$ to arrive at the expansions in \eqref{eq:H_V1_psi}
and \eqref{eq:H_V2_psi}.  Here $H_0$ is the SU($n$)-symmetric
interaction Hamiltonian in \eqref{eq:ints}, $\ket\psi$ is a fully
symmetric state, and $\V_M$ is an $M$-body perturbation of the form
\begin{align}
  \V_M \equiv \sum_{k\in\C_N\p{M}} w_k O_k.
\end{align}
where $O$ is an $M$-body operator that is invariant under arbitrary
permutation of its tensor factors; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
a choice of $M$ distinct spins that $O$ acts on; and $\C_N\p{n}$ is a
choice of $n$ elements from a set of $N$.

@@

We first simplify
\begin{align}
  H_0 \V_1 \ket\psi
  = \sum_k \sum_{p<q} h_{pq} v_k \Pi_{pq} X_k \ket\psi,
  \label{eq:H_V1_psi_start}
\end{align}
where without loss of generality we consider a single-body
perturbation $\V_1$ built out of only one single-body operator $X$.
The sum in \eqref{eq:H_V1_psi_start} has terms with $k\in\set{p,q}$,
and terms with $k\notin\set{p,q}$.  In the case of $k\notin\set{p,q}$,
the permutation operator $\Pi_{pq}$ commutes with $X_k$ and
annihilates on $\ket\psi$, leaving terms at each fixed $k$ of the form
\begin{align}
  \sum_{\substack{p<q\\k\notin\set{p,q}}} h_{pq}
  = \sum_{p<q} h_{pq} - \sum_{\substack{p<q\\k\in\set{p,q}}} h_{pq}
  = E_0 - \f12 \sum_{\substack{p,q\\k\in\set{p,q}}} h_{pq}
  = E_0 - h_k,
\end{align}
where we define $h_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$; and
$h_k \equiv \sum_\ell h_{k\ell}$.  The terms in
\eqref{eq:H_V1_psi_start} with $k\notin\set{p,q}$ are then
\begin{align}
  \sum_{\substack{p<q\\k\notin\set{p,q}}}
  h_{pq} v_k \Pi_{pq} X_k \ket\psi
  = \sum_k \p{E_0 - h_k} v_k X_k \ket\psi
  = E_0 \V_1 \ket\psi - \sum_k h_k v_k X_k \ket\psi
\end{align}
while the terms in \eqref{eq:H_V1_psi_start} with $k\in\set{p,q}$ take
the form
\begin{align}
  \sum_{\substack{p<q\\k\in\set{p,q}}}
  h_{pq} v_k \Pi_{pq} X_k \ket\psi
  = \sum_{k<q} h_{kq} v_k X_q \ket\psi
  + \sum_{p<k} h_{pk} v_k X_p \ket\psi
  = \sum_{k,q} h_{kq} v_k X_q \ket\psi
\end{align}
which implies
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi
  + \sum_p \p{\sum_{q} h_{pq} v_q - h_p v_p}
  X_p \ket\psi.
\end{align}

@@

We wish to simplify
\begin{align}
  H_0 \V_M \ket\psi
  = \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\in\C_N\p{M}}} h_{pq} w_k
  \Pi_{pq} O_k \ket\psi,
  \label{eq:H_VM_psi_start}
\end{align}
where without loss of generality we consider an $M$-body perturbation
$\V_M$ built out of only one $M$-body operator $O$.  The sum in
\eqref{eq:H_VM_psi_start} has terms with $p,q\in k$, terms with
$p,q\notin k$, and terms with only one of $p$ or $q\in k$.  The
permutation operator $\Pi_{pq}$ acts trivially on $O_k\ket\psi$ when
$p,q\in k$ or $p,q\notin k$, leaving terms at each fixed $k$ of the
form
\begin{align}
  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}} h_{pq}
  = \sum_{\p{p,q}\in\C_N\p{2}} h_{pq}
  - \sum_{p\in k} \sum_{\substack{q\in\ZZ_N\\q\notin k}} h_{pq}
  = E_0 - \sum_{p\in k}
  \p{\sum_{q\in\ZZ_N} h_{pq} - \sum_{q\in k} h_{pq}},
\end{align}
where we first split a sum over $\p{p,q}\in\C_N\p{2}$ with the
restriction that $p,q\in k~\t{or}~p,q\notin k$ into an unrestricted
sum minus a remainder, and then similarly split the sum over all spins
$q$ with the restriction that $q\notin k$ into a sum over all $q$
minus the remainder $q\in k$.  We thus find that
\begin{align}
  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}} h_{pq}
  = E_0 + \bar h_k,
  &&
  \bar h_k \equiv 2\sum_{\p{p,q}\subset k} h_{pq} - \sum_{p\in k} h_p,
  &&
  h_p \equiv \sum_q h_{pq},
\end{align}
where the sum over $\p{p,q}\subset k$ denotes a sum over distinct
choices of two elements in $k$.  The terms in
\eqref{eq:H_VM_psi_start} with only one of $p\in k$ or $q\in k$,
meanwhile, are
\begin{align}
  \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{q\in k}
  h_{pq} w_k \Pi_{pq} O_k \ket\psi
  = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  \sum_{j=1}^M h_{p k_j} w_k \Pi_{p k_j} O_k \ket\psi.
\end{align}
Defining
$k_{jp}\equiv\p{k_1,k_2,\cdots,k_{j-1},p,k_{j+1},\cdots,k_M}$, i.e.~an
element of $\C_N\p{M}$ in which the $j$-th number $k_j$ has been
replaced by $p$, we can simplify
\begin{align}
  \Pi_{p k_j} O_k \ket\psi = O_{k_{jp}}\ket\psi,
\end{align}
and switch the labels for $p$ and $k_j$ to find that
\begin{align}
  \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{q\in k}
  h_{pq} w_k  \Pi_{pq} O_k \ket\psi
  = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  \sum_{j=1}^M h_{pk_j} w_{k_{jp}} O_k \ket\psi.
\end{align}
Defining
\begin{align}
  \p{h\circ w}_k
  \equiv \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{j=1}^M
  h_{pk_j} w_{k_{jp}}
  = \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  \p{h_{pk_1} w_{pk_2k_3\cdots k_M}
    + h_{pk_2} w_{k_1pk_3\cdots k_M}
    + \cdots + h_{pk_M} w_{k_1k_2\cdots k_{M-1} p}},
\end{align}
we thus find that
\begin{align}
  H_0 \V_M \ket\psi
  = E_0 \V_M \ket\psi + \sum_{k\in\C_N\p{M}}
  \sp{\bar h_k w_k + \p{h\circ w}_k} O_k \ket\psi.
\end{align}
If $M=1$, then $\bar h_k=-h_k$ and
$\p{h\circ w}_k=\sum_\ell h_{k\ell}w_\ell$, so
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi + \sum_{k\in\ZZ_N}
  \p{\sum_\ell h_{k\ell} w_\ell - h_k w_k} O_k \ket\psi.
\end{align}
If $M=2$, then
$\p{h\circ w}_{pq}=\sum_k\p{h_{kp} w_{kq}+h_{kq}w_{pk}}$, so
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi + \sum_{p<q}
  \sp{\bar h_{pq} w_{pq} + \sum_k\p{h_{kp} w_{kq} + h_{kq} w_{pk}}}
  O_k \ket\psi,
\end{align}
where $\bar h_{pq}=2h_{pq}-h_p-h_q$ with $h_p\equiv\sum_qh_{pq}$.

@@

Given a system of $N$ spins, we wish to project a product of
multi-body operators onto the fully symmetric manifold.  For $p$
multi-body operators, the projection of such a product takes the form
\begin{align}
  \P_0 \sp{\prod_{j\in\ZZ_p}
    \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}} \P_0,
  \label{eq:sym_prod_proj}
\end{align}
where $\P_0$ is a projector onto the fully symmetric manifold;
$k\in\C_N\p{M_j}$ is a choice of $M_j$ distinct spins; $O_j\p{k}$
denotes the action of an $M_j$-body operator on spins $k$ that is
invariant under permutations of $k$; and $w_j\p{k}$ is a scalar
coefficient for $O_j\p{k}$.

>>>>

%%% figures
\usepackage{graphicx} % for figures
\graphicspath{{./figures/}} % set path for all figures
\usepackage[export]{adjustbox} % for vertical alignment in math
\newcommand{\diagram}[1]
{\,\includegraphics[valign=c]{diagrams/#1.pdf}\,}

@@

We wish to determine the effective dynamics induced on the
ground-state manifold $\M_0$ of $H_0$ by weak single- and two-body
perturbations of the form
\begin{align}
  \V_1 \equiv \sum_X \sum_p v_{Xp} X_p = \sum_X \v v_X\c\v X,
  &&
  \V_2 \equiv \sum_O \sum_{p<q} w_{Opq} O_{pq}
  = \sum_O \v{\m w}_O\c \v{\m O}
  \label{eq:perturbations}
\end{align}
where $p,q$ index individual spins; $v_{Xp}$ and $w_{Opq}$ are real
numbers; $X$ is a trace-zero single-spin operator on $\H_n$; $O$ is a
trace-zero two-spin operator on $\H_n\otimes\H_n$ that obeys pair-wise
permutational symmetry, i.e.~$O_{pq}=O_{qp}$; $\v v_X$ and $\v X$ are
$N$-component vectors defined by
\begin{align}
  \v v_X \equiv \sum_p v_{Xp} \ket{p},
  &&
  \v X \equiv \sum_p X_p \ket{p};
\end{align}
and finally $\v{\m w_O}$ and $\v{\m O}$ are ${N \choose 2}$-component
vectors defined by
\begin{align}
  \v{\m w}_O \equiv \sum_{p<q} w_{Opq} \ket{\set{p,q}},
  &&
  \v{\m O} \equiv \sum_{p<q} O_{pq} \ket{\set{p,q}},
\end{align}
where the basis vectors $\ket{\set{p,q}}$ are labeled by a choice of
two distinct spins $p,q$.

@@

If we {\it construct} a perturbation $\V_1^\Delta$ with coefficients
$\v v_X^\Delta$ that satisfy \eqref{eq:cond_1} with eigenvalue
$\Delta$, then
\begin{align}
  H_0 \V_1^\Delta \ket\psi = \p{E_0 + \Delta} \V_1^\Delta \ket\psi.
\end{align}
Interestingly, the energy of the state $\V_1^\Delta\ket\psi$ depends
only on the coefficients $\v v_X^\Delta$, and is entirely independent
of the state $\ket\psi$ or the choice of a trace-zero single-spin
operators $X$ used to build $\V_1^\Delta$.  Finding operators
$\V_1^\Delta$ that generate eigenvectors of the interaction
Hamiltonian $H_0$ when they act on fully symmetric states $\ket\psi$
thus reduces to finding eigenvectors of the interaction matrix
$\m h-\diag\v h$.

@@

We now return to the task of computing the second-order effective
Hamiltonian $H_1^{(2)}$.  Any coefficient vector $\v v_X$ can be
expanded into its projections $\v v_X^\Delta$ onto the eigenspace of
vectors satisfying \eqref{eq:cond_1} with eigenvalue $\Delta$,
i.e.~$\v v_X = \sum_\Delta \v v_X^\Delta$, which also allows us to
expand
\begin{align}
  \V_1 = \sum_\Delta \V_1^\Delta,
  &&
  \V_1^\Delta \equiv \sum_X \v v_X^\Delta \c \v X,
\end{align}
where each operator $\V_1^\Delta$ generates a state with definite
interaction energy $\Delta$ above that of the fully symmetric manifold
$\M_0$.  We can therefore simplify
\begin{align}
  H_1^{(2)}
  = - \P_0 \V_1 \E \V_1 \P_0
  = - \sum_{\Delta\ne0} \f1{\Delta}
  \P_0 \V_1^\Delta \P_\Delta \V_1^\Delta \P_0
  = - \sum_{\Delta\ne0} \f1{\Delta} \P_0 \p{\V_1^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_1^\Delta}^2 \P_0$ is worked out in
Appendix \ref{sec:PXYP}.  In total, we find
\begin{align}
  H_1^{(2)}
  \EQFS \f1{N\p{N-1}} \sum_{X,Y} \sum_{\Delta\ne0}
  \f{\v v_X^\Delta\c\v v_Y^\Delta}{\Delta}
  \p{\col{X}\,\col{Y} - N \col{XY}},
  \label{eq:H_1_2}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$; $\v v_X^\Delta$ is the projection of
$\v v_X$ onto the $\Delta$-eigenspace of the matrix $\m h-\diag\v h$;
and for any single-spin operator $Z$, we define the collective
operator
\begin{align}
  \col{Z} \equiv \sum_p Z_p.
\end{align}

@@

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
\eqref{eq:perturbations} is simply the restriction of $\V_2$ onto
$\M_0$:
\begin{align}
  H_2^{(1)} = \sum_O \sum_{p<q} w_{Opq} \P_0 O_{pq} \P_0
  \EQFS \sum_O \sum_{p<q} w_{Opq} O_{1,2}
  \EQFS \sum_O \bar w_O \col{O},
\end{align}
where
\begin{align}
  \bar w_O \equiv {N \choose 2}^{-1} \sum_{p<q} w_{Opq},
  &&
  \col{O} \equiv \sum_{p<q} O_{pq},
\end{align}
are an average of $w_{Opq}$ and a collective version of the two-spin
operator $O_{pq}$.  If the two-spin operator $O=X\otimes Y$, then
\begin{align}
  H_2^{(1)} \EQFS \f12 \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
  \p{\col{X}\,\col{Y} - \col{XY}}.
\end{align}
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_VM_psi})
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \sum_O \sum_{p<q} \sp{\bar h_{pq} w_{Opq}
    + \sum_k \p{h_{kp} w_{Okq} w_{Okq} + h_{kq} w_{Opk}}}
  O_{pq} \ket\psi,
  \label{eq:H_V2_psi}
\end{align}
where to keep notation compact, we define
\begin{align}
  \bar h_{pq} \equiv 2 h_{pq} - h_p - h_q,
  &&
  h_p \equiv \sum_q h_{pq},
\end{align}
and we define $w_{Opq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$.  We thus find that the
vector $\V_2^\Delta\ket\psi$ (indexed by $\Delta$) is an eigenvector
of the interaction Hamiltonian $H_0$ if its coefficients
$\v{\m w}_O^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \check{\m h} \c \v{\m w} = \Delta \v{\m w},
  \label{eq:cond_2}
\end{align}
where
\begin{align}
  \check{\m h} \equiv \sum_{p<q} \bar h_{pq} \op{\set{p,q}}
  + \sum_{p<q} \ket{\set{p,q}} \sum_{k\notin\set{p,q}}
  \p{h_{kp} \bra{\set{k,q}} + h_{kq} \bra{\set{p,k}}}.
  \label{eq:h_super_mat}
\end{align}
Solving the general eigenvalue problem in \eqref{eq:cond_2} requires
diagonalizing the ${N \choose 2}\times{N \choose 2}\sim N^2\times N^2$
matrix $\check{\m h}$.  When both he interaction Hamiltonian $H_0$
translationally invariant, however, this eigenvalue problem can be
reduced to that of diagonalizing an $\sim N\times N$ matrix; we
discuss this reduction in Appendix \ref{sec:trans_inv}.

@@

In this section, we simplify the products of the form
$H_0\V_M\ket\psi$ to arrive at the expansions in \eqref{eq:H_V1_psi}
and \eqref{eq:H_V2_psi}.  Here $H_0$ is the SU($n$)-symmetric
interaction Hamiltonian in \eqref{eq:ints}, $\ket\psi$ is a fully
symmetric state, and $\V_M$ is an $M$-body perturbation of the form
\begin{align}
  \V_M \equiv \sum_{k\in\C_N\p{M}} w_k O_k.
\end{align}
where $O$ is an $M$-spin operator that is invariant under arbitrary
permutation of its tensor factors; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
a choice of $M$ distinct spins that $O$ acts on; and $\C_N\p{n}$ is a
choice of $n$ elements from a set of $N$.

@@

We first simplify
\begin{align}
  H_0 \V_1 \ket\psi
  = \sum_k \sum_{p<q} h_{pq} v_k \Pi_{pq} X_k \ket\psi,
  \label{eq:H_V1_psi_start}
\end{align}
where without loss of generality we consider a single-body
perturbation $\V_1$ built out of only one single-spin operator $X$.
The sum in \eqref{eq:H_V1_psi_start} has terms with $k\in\set{p,q}$,
and terms with $k\notin\set{p,q}$.  In the case of $k\notin\set{p,q}$,
the permutation operator $\Pi_{pq}$ commutes with $X_k$ and
annihilates on $\ket\psi$, leaving terms at each fixed $k$ of the form
\begin{align}
  \sum_{\substack{p<q\\k\notin\set{p,q}}} h_{pq}
  = \sum_{p<q} h_{pq} - \sum_{\substack{p<q\\k\in\set{p,q}}} h_{pq}
  = E_0 - \f12 \sum_{\substack{p,q\\k\in\set{p,q}}} h_{pq}
  = E_0 - h_k,
\end{align}
where we define $h_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$; and
$h_k \equiv \sum_\ell h_{k\ell}$.  The terms in
\eqref{eq:H_V1_psi_start} with $k\notin\set{p,q}$ are then
\begin{align}
  \sum_{\substack{p<q\\k\notin\set{p,q}}}
  h_{pq} v_k \Pi_{pq} X_k \ket\psi
  = \sum_k \p{E_0 - h_k} v_k X_k \ket\psi
  = E_0 \V_1 \ket\psi - \sum_k h_k v_k X_k \ket\psi
\end{align}
while the terms in \eqref{eq:H_V1_psi_start} with $k\in\set{p,q}$ take
the form
\begin{align}
  \sum_{\substack{p<q\\k\in\set{p,q}}}
  h_{pq} v_k \Pi_{pq} X_k \ket\psi
  = \sum_{k<q} h_{kq} v_k X_q \ket\psi
  + \sum_{p<k} h_{pk} v_k X_p \ket\psi
  = \sum_{k,q} h_{kq} v_k X_q \ket\psi
\end{align}
which implies
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi
  + \sum_p \p{\sum_{q} h_{pq} v_q - h_p v_p}
  X_p \ket\psi.
\end{align}

@@

We wish to simplify
\begin{align}
  H_0 \V_M \ket\psi
  = \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\in\C_N\p{M}}} h_{pq} w_k
  \Pi_{pq} O_k \ket\psi,
  \label{eq:H_VM_psi_start}
\end{align}
where without loss of generality we consider an $M$-body perturbation
$\V_M$ built out of only one $M$-spin operator $O$.  The sum in
\eqref{eq:H_VM_psi_start} has terms with $p,q\in k$, terms with
$p,q\notin k$, and terms with only one of $p$ or $q\in k$.  The
permutation operator $\Pi_{pq}$ acts trivially on $O_k\ket\psi$ when
$p,q\in k$ or $p,q\notin k$, leaving terms at each fixed $k$ of the
form
\begin{align}
  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}} h_{pq}
  = \sum_{\p{p,q}\in\C_N\p{2}} h_{pq}
  - \sum_{p\in k} \sum_{\substack{q\in\ZZ_N\\q\notin k}} h_{pq}
  = E_0 - \sum_{p\in k}
  \p{\sum_{q\in\ZZ_N} h_{pq} - \sum_{q\in k} h_{pq}},
\end{align}
where we first split a sum over $\p{p,q}\in\C_N\p{2}$ with the
restriction that $p,q\in k~\t{or}~p,q\notin k$ into an unrestricted
sum minus a remainder, and then similarly split the sum over all spins
$q$ with the restriction that $q\notin k$ into a sum over all $q$
minus the remainder $q\in k$.  We thus find that
\begin{align}
  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}} h_{pq}
  = E_0 + \bar h_k,
  &&
  \bar h_k \equiv 2\sum_{\p{p,q}\subset k} h_{pq} - \sum_{p\in k} h_p,
  &&
  h_p \equiv \sum_q h_{pq},
\end{align}
where the sum over $\p{p,q}\subset k$ denotes a sum over distinct
choices of two elements in $k$.  The terms in
\eqref{eq:H_VM_psi_start} with only one of $p\in k$ or $q\in k$,
meanwhile, are
\begin{align}
  \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{q\in k}
  h_{pq} w_k \Pi_{pq} O_k \ket\psi
  = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  \sum_{j=1}^M h_{p k_j} w_k \Pi_{p k_j} O_k \ket\psi.
\end{align}
Defining
$k_{jp}\equiv\p{k_1,k_2,\cdots,k_{j-1},p,k_{j+1},\cdots,k_M}$, i.e.~an
element of $\C_N\p{M}$ in which the $j$-th number $k_j$ has been
replaced by $p$, we can simplify
\begin{align}
  \Pi_{p k_j} O_k \ket\psi = O_{k_{jp}}\ket\psi,
\end{align}
and switch the labels for $p$ and $k_j$ to find that
\begin{align}
  \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{q\in k}
  h_{pq} w_k  \Pi_{pq} O_k \ket\psi
  = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  \sum_{j=1}^M h_{pk_j} w_{k_{jp}} O_k \ket\psi.
\end{align}
Defining
\begin{align}
  \p{h\circ w}_k
  \equiv \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{j=1}^M
  h_{pk_j} w_{k_{jp}}
  = \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  \p{h_{pk_1} w_{pk_2k_3\cdots k_M}
    + h_{pk_2} w_{k_1pk_3\cdots k_M}
    + \cdots + h_{pk_M} w_{k_1k_2\cdots k_{M-1} p}},
\end{align}
we thus find that
\begin{align}
  H_0 \V_M \ket\psi
  = E_0 \V_M \ket\psi + \sum_{k\in\C_N\p{M}}
  \sp{\bar h_k w_k + \p{h\circ w}_k} O_k \ket\psi.
\end{align}
If $M=1$, then $\bar h_k=-h_k$ and
$\p{h\circ w}_k=\sum_\ell h_{k\ell}w_\ell$, so
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi + \sum_{k\in\ZZ_N}
  \p{\sum_\ell h_{k\ell} w_\ell - h_k w_k} O_k \ket\psi.
\end{align}
If $M=2$, then
$\p{h\circ w}_{pq}=\sum_k\p{h_{kp} w_{kq}+h_{kq}w_{pk}}$, so
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi + \sum_{p<q}
  \sp{\bar h_{pq} w_{pq} + \sum_k\p{h_{kp} w_{kq} + h_{kq} w_{pk}}}
  O_k \ket\psi,
\end{align}
where $\bar h_{pq}=2h_{pq}-h_p-h_q$ with $h_p\equiv\sum_qh_{pq}$.

@@

\subsection{The general case}

Given a system of $N$ spins, we wish to project a product of
multi-body operators onto the fully symmetric manifold.  For $p$
multi-body operators, the projection of such a product takes the form
\begin{align}
  \P_0 \sp{\prod_{j\in\ZZ_p}
    \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}} \P_0,
  \label{eq:sym_prod_proj}
\end{align}
where $\P_0$ is a projector onto the fully symmetric manifold;
$k\in\C_N\p{M_j}$ is a choice of $M_j$ distinct spins; $O_j\p{k}$
denotes the action of an $M_j$-spin operator on spins $k$ that is
invariant under permutations of $k$; and $w_j\p{k}$ is a scalar
coefficient for $O_j\p{k}$.
